---
title: "TidyR"
output: 
  bookdown::html_document2:
      base_format: rmarkdown::html_vignette
      fig_caption: yes
      number_sections: no
vignette: >
  %\VignetteIndexEntry{TidyR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7,fig.height = 5,
                      fig.align = "center", message = FALSE,
                      collapse = TRUE, comment = "#>" )

library(TidyR)     
library(janitor)    
library(ggplot2)
```

## Introduction

__TidyR__ is an R package which tidies data, saves this tidied dataset and plots it. There are three functions provided:

  Function 1 : The `tidy_data` function makes column names the same and removes any NA values.
  
  Function 2: The `load_data` function saves the processed dataset and makes available as part of this R package
  
  Function 3: The `plot.tidy_data` function generates a plot for the inputted x vs y based on `ggplot2`.


### Installation

__TidyR__ will run in Windows, Mac OS X, or Linux. To install it you first need to install [R](https://cran.r-project.org/). Installing [RStudio](https://www.rstudio.com/) as a nice desktop environment for using R is also recommended.

In R you can type the following into the Console tab:

```{r, eval=FALSE}
install.packages('devtools')
devtools::install_github("erinweadick/TidyR")
```

To install the latest development version of the package from the __TidyR__ [GitHub page](https://github.com/erinweadick/TidyR). 

In either case, if you then type:

```{r, eval=FALSE}
library(TidyR)
```

it will load in all the aforementioned __TidyR__ functions.

### Load data
```{r data-prep}
library(TidyR)

data(raw_data)


head(raw_data)
```

## tidy_data

This function converts column names to a specified case style (e.g., snake_case, UpperCamelCase).
It removes rows with NA values.
It ensures compatibility with other functions in the package.

```{r}
tidy_data <- function(data, case = "snake") {

  if (!is.data.frame(data)) {
    stop("Input must be a data frame")
  }

  tidied_data <- data %>%
    na.omit() %>%
    clean_names(case = case)

  class(tidied_data) <- c("tidy_data", class(tidied_data))
  tidied_data
}

tidied_data <- tidy_data(raw_data, case = "snake")
```


## load_data

This function saves the processed dataset as an R object, allowing it to be used seamlessly within the package. It is saved using the 'usethis :: use_data' function.

```{r}
load_data <- function(data){
  usethis::use_data(data, compress = "xz", overwrite = TRUE)
}
```


## plot.tidy_data

This final function generates a scatter plot. The function checks if var_x and var_y are provided. If either is missing, an error message is displayed: "Both 'var_x' and 'var_y' must be specified via ...".
Our `geom_point` adds scatter plot points, our `geom_smooth` adds a linear regression line without the confidence interval.
`theme_minimal` produces a clean, minimalistic look.

```{r}
plot.tidy_data <- function(x, y = NULL, ...) {
  dots <- list(...)
  var_x <- dots$var_x
  var_y <- dots$var_y

  if (is.null(var_x) || is.null(var_y)) {
    stop("Both 'var_x' and 'var_y' must be specified via ...")
  }

  ggplot2::ggplot(x, ggplot2::aes(x = .data[[var_x]], y = .data[[var_y]])) +
    ggplot2::geom_point(color = "steelblue", size = 3, alpha = 0.8) +
    ggplot2::geom_smooth(
      method = "lm", color = "darkred", se = FALSE,
      linetype = "dashed", linewidth = 1
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(
        hjust = 0.5, face = "bold",
        color = "darkblue", size = 16
      ),
      panel.grid.major = ggplot2::element_line(
        color = "lightgray",
        linetype = "dotted"
      ),
      panel.grid.minor = ggplot2::element_blank()
    )
}
#Example
plot(tidied_data,var_x = "chlorophyll",var_y = "spad")
```

The output of the plot.tidy_data function is a scatter plot using ggplot2. This includes a dashed, red linear regression line, allowing users to visualize relationships between two variables in the tidied dataset. The points of the scatterplot are a 'steelblue' colour. This is displayed over a 'lightgrey' panel grid.
